---
title: "project.Rmd"
author: "RAGINI"
date: "19/08/2020"
output: html_document
---
##Step 1
##Code for reading in the dataset and/or processing the data
```{r, echo = TRUE}
setwd("C:/Users/Ragini Banala/Downloads")
activity<-read.csv("activity.csv")
```
Exploring the basics of this data
```{r, echo = TRUE}
dim(activity)
names(activity)
head(activity)
str(activity)
#total number of missing data
sum(is.na(activity$steps))/dim(activity)[[1]]
#transforming the date column into date format using lubridate
library(lubridate)
activity$date<-ymd(activity$date)
length(unique(activity$date))
```
##Step 2
##Histogram of the total number of steps taken each day

```{r, echo = TRUE}
StepsPerDay <- aggregate(activity$steps, list(activity$date), FUN=sum)
colnames(StepsPerDay) <- c("Date", "Steps")
StepsPerDay
library(ggplot2)
png("plot1.png")
g <- ggplot(StepsPerDay, aes(Steps))
g+geom_histogram(boundary=0, binwidth=2500, col="darkgreen", fill="lightgreen")+ggtitle("Histogram of steps per day")+xlab("Steps")+ylab("Frequency")+theme(plot.title = element_text(face="bold", size=12))+scale_x_continuous(breaks=seq(0,25000,2500))+scale_y_continuous(breaks=seq(0,18,2))
dev.off()
```
##Step 3
##Mean and median number of steps taken each day

```{r, echo = TRUE}
#mean
mean(StepsPerDay$Steps, na.rm=TRUE)

#Median
median(StepsPerDay$Steps, na.rm=TRUE)

```
##Step 4
##Time series plot of the average number of steps taken
```{r, echo = TRUE}
# create table with steps per time
StepsPerTime <- aggregate(steps~interval,data=activity,FUN=mean,na.action=na.omit)
# variable time (more comprensible for the graph axis)
StepsPerTime$time <- StepsPerTime$interval/100
# draw the line plot
png("plot2.png")
h <- ggplot(StepsPerTime, aes(time, steps))
h+geom_line(col="brown")+ggtitle("Average steps per time interval")+xlab("Time")+ylab("Steps")+theme(plot.title = element_text(face="bold", size=12))
dev.off()
```
##Step 5
##The 5-minute interval that, on average, contains the maximum number of steps

```{r, echo = TRUE}
# table for dplyr
library(dplyr)
ST <- tbl_df(StepsPerTime)
# find the column
ST %>% select(time, steps) %>% filter(steps==max(ST$steps))
```
##Imputing missing values
###First, we calculate and report the total number of missing values in the dataset (i.e. the total number of rows with NAs):
```{r}
sum(is.na(activity$steps))
#So the original data set has 2304 rows with missing data.

#We use a simple strategy for filling in all of the missing values in the dataset. If a 5-minute interval has missing value, we use the mean for that 5-minute interval.

#We create a new data frame df_impute that is equal to the original dataset but with the missing data filled in (using mean for that interval for imputation):
activity$date <- as.Date(activity$date)
df_ign <- subset(activity, !is.na(activity$steps))
df_impute <- activity
ndx <- is.na(df_impute$steps)
int_avg <- tapply(df_ign$steps, df_ign$interval, mean, na.rm=TRUE, simplify=T)

df_impute$steps[ndx] <- int_avg[as.character(df_impute$interval[ndx])]

new_dailysum <- tapply(df_impute$steps, df_impute$date, sum, na.rm=TRUE, simplify=T)
png("plot3.png")
hist(x=new_dailysum,
     col="red",
     breaks=20,
     xlab="daily steps",
     ylab="frequency",
     main="The distribution of daily total (with missing data imputed)")
dev.off()
mean(new_dailysum)

median(new_dailysum)
```

##Are there differences in activity patterns between weekdays and weekends?
###First we create a new factor variable “wk” in the dataset with two levels - “weekday” and “weekend” indicating whether a given date is a weekday or weekend day.
```{r}

# helper function to decide if a day is a week day or not
is_weekday <- function(d) {
    wd <- weekdays(d)
    ifelse (wd == "Saturday" | wd == "Sunday", "weekend", "weekday")
}

wx <- sapply(df_impute$date, is_weekday)
df_impute$wk <- as.factor(wx)
head(df_impute)

wk_df <- aggregate(steps ~ wk+interval, data=df_impute, FUN=mean)
#Next we make a panel plot containing a time series plot (i.e. type = “l”) of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis).

library(lattice)
png("plot4.png")
xyplot(steps ~ interval | factor(wk),
       layout = c(1, 2),
       xlab="Interval",
       ylab="Number of steps",
       type="l",
       lty=1,
       data=wk_df)
dev.off()
```

